---
title: "Correlation Tables"
author: "Matthew Peverill"
date: "12/12/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(pander)
library(sdlabFunctions)

#Set up pander
panderOptions("digits",2)
panderOptions("table.split.table",150)
panderOptions("table.split.cells",20)
```

Let's explore a few options for correlation tables.

# apaTables

apaTables::apa.cor.table does a lot, but is a bit inflexible. You cannot turn off the confidence intervals or the explanatory text. It will print a doc file for you if you'd like, which I'd recommend if you want to use it.

```{r}
pander(apaTables::apa.cor.table(mtcars,filename="apacortable.doc"))
```

# Custom table with Hmisc

A more flexible method is provided by sdcorr from sdlabFunctions. Here is the source:

`r pander(sdcorr)`

```{r}
cortable<-sdcorr(mtcars)
pander(cortable)
```

As elsewhere, I recommend saving a csv file and editing in excel (once you establish your formatting and labels in an xlsx, you can just copy the raw values over as needed).

```{r}
write.csv(cortable,file="cortable.csv")
```
